---
title: "find_Ns_Np"
author: "Qianqian Du"
output: html_document
---

# uplode the packages
```{r}
library(apsimx)
library(tidyverse)
library(ggplot2)
library(data.table)
```

# run a simulation through R

```{r}
# load a apsim file

extd.dir = "/Users/qianqiandu/Desktop/12_Nitrogen_and_Carbon_Cycling"

inspect_apsimx("Nitrogen_balance.apsimx", src.dir = extd.dir, node = "Soil")

sim <- apsimx("Nitrogen_balance.apsimx", src.dir = extd.dir, value = "Report") 

class(sim$Date)

result <- sim %>% 
  mutate(year = as.numeric(format(sim$Date, "%Y"))) %>% 
  mutate(month = as.numeric(format(sim$Date, "%m"))) %>% 
  mutate(day = as.numeric(format(sim$Date, "%d")))

names(sim)
  
  
# See surface N
ggplot(data = result)+
  geom_point(aes(x = Clock.Today, y = SurfaceN))


result_analysis <- result %>% 
  data.table() %>% 
  .[, .(year_surface_N = mean(SurfaceN)), by = c("year", "month")] %>% 
  .[, year_month := paste0(year, month)]


ggplot(data = result_analysis)+
  geom_point(aes(x = year_month, y = year_surface_N))


# See yield and Surface N
scale = 40

ggplot(data = result)+
  geom_point(aes(x = Clock.Today, y = CropYield))+
  geom_point(aes(x = Clock.Today, y = SurfaceN*scale), color = "red")+
  scale_y_continuous(
    name = "yield",
    sec.axis = sec_axis(~.*1/scale, name="Surface N")
  )
```


I also run by a different soil type, the surface_N is very different.
run by different N amount, for different years.
seq(100, 200, by = 20), isolate the impact of weather. 

Next:

+ Figure out N_s and N_p, which two variables actually represent the N_s and N_p, then see how vary they are.
+ Find a point's historical daily weather data, finding a function in R to generate its weather distribution.
+ Run APSIM on the hypothetical weather data for that point.
+ Solve the ex-ante and ex-post for maximization.
