---
title: "Consistent N rate trials over years"
author: "Qianqian Du"
date: "2024-03-04"
output: html_document
---

# Objectives:
+ run on-farm experiments over years using the same trial design each year, assume homogeneous environment.

# uplode the packages
```{r, message=FALSE, warning=FALSE}
# update the apsimx package as the latest version
# library(remotes)
# remotes::install_github('femiguez/apsimx')

library(apsimx)
library(tidyverse)
library(ggplot2)
library(data.table)
library(oce)
library(lubridate)
library(daymetr)

``` 


```{r}

### ============ set path to the file ============ ###
extd.dir = "/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/consistent_rate_trials"

write_dir = "/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/consistent_rate_trials"


###========== Check parameters of the file =========###
inspect_apsimx("consisten_rate_trials.apsimx", src.dir = extd.dir, node = "Clock")


edit_apsimx("consisten_rate_trials.apsimx", src.dir = extd.dir,
              wrt.dir = write_dir, node = "Clock",
              parm = "Start", value = "01/01/2015")

edit_apsimx("consisten_rate_trials-edited.apsimx", src.dir = extd.dir,
              wrt.dir = write_dir, node = "Clock",
              parm = "End", value = "12/31/2021", overwrite = TRUE)

inspect_apsimx("consisten_rate_trials-edited.apsimx", src.dir = extd.dir, node = "Clock")


inspect_apsimx("consisten_rate_trials-edited.apsimx", src.dir = extd.dir, node = "Weather")

edit_apsimx("consisten_rate_trials-edited.apsimx", src.dir = extd.dir,
              wrt.dir = write_dir, node = "Weather",
              value = "fake_weather_dry.met", overwrite = TRUE)

inspect_apsimx("consisten_rate_trials-edited.apsimx", src.dir = extd.dir, node = "Weather")


```


# See the result
```{r}

analysis_trials_data <- readRDS(paste0(write_dir, "/consistent_rate_trials.rds"))

# Take a look at the NO3 levels from each N rate application

NO3_by_N <- ggplot(data = analysis_trials_data)+
  geom_line(aes(x = Date, y = NO3, color = applied_N))+
  theme_bw()

## The figure of NO3 by years 

NO3_by_N


# Find the max yield level in each year in each level of N application

analysis_result <- analysis_trials_data %>% 
  data.table() %>% 
  .[, .(max_yield = max(CropYield)), by = c("year", "applied_N")]

ggplot(data = analysis_result)+
  geom_point(aes(x = applied_N, y = max_yield, color = as.factor(year)))+
  geom_line(aes(x = applied_N, y = max_yield, color = as.factor(year)))

```

```{r}
max_yield <- analysis_trials_data %>% 
  .[, .(maxYield = max(CropYield)), by = c("year")]

ggplot(data = max_yield)+
  geom_point(aes(x = year, y = maxYield))


#--- download weather data in the field location from daymat ---#
daymet_data <- download_daymet("Ames",
                               lat = 42.0308, 
                               lon = -93.6319, 
                               start = 2004, 
                               end = 2004,
                               path = "/Users/qianqiandu/Desktop",
                               internal = FALSE)


read_csv <- read_csv(("/Users/qianqiandu/Desktop/Ames_2004_2004.csv"), skip = 6) %>% 
  data.table() %>% 
  setnames("prcp (mm/day)", "prcp", skip_absent=TRUE) %>% 
  setnames('tmax (deg c)', "tmax", skip_absent=TRUE) %>% 
  setnames('tmin (deg c)', "tmim", skip_absent=TRUE) %>% 
  # .[, week := paste0("week_", sprintf("%02d", ceiling(yday / 7)))] %>% 
  .[, mean_temp := (tmax + tmim)/2] %>%
  .[, date := yday - 1 + ymd(paste0(year, "-01-01"))] %>%
  .[, month := month(date)]


```

```{r}
#------- get weekly average precipitation and temperature data -------#
weather_weekly <- read_csv %>% 
  group_by(year, week) %>% 
  summarise(week_prcp = mean(prcp), week_temp = mean(mean_temp)) %>% 
  data.table()


#------- get yearly average precipitation -------#
weather_result_year <- read_csv %>% 
  data.table() %>% 
  .[year<=2020 & year >= 2005, .(month_precip = sum(prcp)), by = c("year")] %>% 
  # .[, .(mean_precip = mean(month_precip)), by = "month"] %>% 
  ##---- convert precip from mm to inch ----##
  .[, year_precip_in := month_precip* 0.0393701]



### The figure of max yield and precipitation ###
ggplot()+
  geom_line(aes(x = year, y = max_yield, color = as.factor(applied_N)), data = analysis_result)+
  geom_line(aes(x = year, y = year_precip_in*10), data = weather_result_year)+
  # Custom the Y scales:
  scale_y_continuous(
    
    # Features of the first axis
    name = "yield",
    
    # Add a second axis and specify its features
    sec.axis = sec_axis(trans=~.*0.1, name = "precip")
  )+
  theme_bw()
  
```



# change the N rates in Operations through R

```{r}
simulation_result <- data.frame()
simulation_wet_weather <- data.frame()

# length(N_rates[[1]])
```

```{r}

###=============================================================###
#####======== change the N rates in Operations through R =========#
###=============================================================###

#### Editing just one operation, need to change all three ####

N_rates <- list(seq(0, 270, by = 30))

for (i in 1:length(N_rates[[1]])) {

N = as.character(N_rates[[1]][i])

new.date <- "2015-05-05"
# new.action <- '[Fertiliser].Apply(Amount: 0, Type: Fertiliser.Types.UAN_N, Depth: 30)'
new.action <- paste0("[Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")

# new.line <- '2015-05-05 [Fertiliser].Apply(Amount: 0, Type: Fertiliser.Types.UAN_N, Depth: 30)'
new_line <- paste0("2015-05-05 [Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(3, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(3, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(3, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(3, 'Date'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(3, 'Action'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(3, 'Line'))


##--------------- change the 2nd year's fertilizer amount -------------##
new.date <- "2016-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2016-05-05 [Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(11, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(11, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(11, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(11, 'Date'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(11, 'Action'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(11, 'Line'))



##--------------- change the 3rd year's fertilizer amount -------------##
new.date <- "2017-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2017-05-05 [Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(19, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(19, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(19, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(19, 'Date'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(19, 'Action'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(19, 'Line'))


##--------------- change the 4th year's fertilizer amount -------------##
new.date <- "2018-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2018-05-05 [Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(27, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(27, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(27, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(27, 'Date'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(27, 'Action'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(27, 'Line'))


##--------------- change the 5th year's fertilizer amount -------------##
new.date <- "2019-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2019-05-05 [Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(35, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(35, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(35, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(35, 'Date'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(35, 'Action'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(35, 'Line'))



##--------------- change the 6th year's fertilizer amount -------------##
new.date <- "2020-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2020-05-05 [Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(43, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(43, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(43, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(43, 'Date'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(43, 'Action'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(43, 'Line'))


##--------------- change the 7th year's fertilizer amount -------------##
new.date <- "2021-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2021-05-05 [Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(51, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(51, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("consisten_rate_trials-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(51, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(51, 'Date'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(51, 'Action'))

inspect_apsimx("consisten_rate_trials-edited.apsimx",
               node = "Operations",
               parm = list(51, 'Line'))



##--------------- change the 8th year's fertilizer amount -------------##
# new.date <- "2022-05-05"
# new.action <- paste0("[Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
# new_line <- paste0("2022-05-05 [Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
# 
# edit_apsimx("consisten_rate_trials-edited.apsimx",
#             src.dir = write_dir,
#             wrt.dir = write_dir,
#             node = "Operations",
#             parm = list(59, 'Date'),
#             value = new.date,
#             overwrite = TRUE)
# 
# 
# edit_apsimx("consisten_rate_trials-edited.apsimx",
#             src.dir = write_dir,
#             wrt.dir = write_dir,
#             node = "Operations",
#             parm = list(59, 'Action'),
#             value = new.action,
#             overwrite = TRUE)
# 
# 
# edit_apsimx("consisten_rate_trials-edited.apsimx",
#             src.dir = write_dir,
#             wrt.dir = write_dir,
#             node = "Operations",
#             parm = list(59, 'Line'),
#             value = new.action,
#             overwrite = TRUE)
# 
# inspect_apsimx("consisten_rate_trials-edited.apsimx",
#                node = "Operations",
#                parm = list(59, 'Date'))
# 
# inspect_apsimx("consisten_rate_trials-edited.apsimx",
#                node = "Operations",
#                parm = list(59, 'Action'))
# 
# inspect_apsimx("consisten_rate_trials-edited.apsimx",
#                node = "Operations",
#                parm = list(59, 'Line'))


##--------------- change the 9th year's fertilizer amount -------------##
# new.date <- "2023-05-05"
# new.action <- paste0("[Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
# new_line <- paste0("2023-05-05 [Fertiliser].Apply(Amount: ",N,", Type: Fertiliser.Types.UAN_N, Depth: 30)")
# 
# edit_apsimx("consisten_rate_trials-edited.apsimx",
#             src.dir = write_dir,
#             wrt.dir = write_dir,
#             node = "Operations",
#             parm = list(67, 'Date'),
#             value = new.date,
#             overwrite = TRUE)
# 
# 
# edit_apsimx("consisten_rate_trials-edited.apsimx",
#             src.dir = write_dir,
#             wrt.dir = write_dir,
#             node = "Operations",
#             parm = list(67, 'Action'),
#             value = new.action,
#             overwrite = TRUE)
# 
# 
# edit_apsimx("consisten_rate_trials-edited.apsimx",
#             src.dir = write_dir,
#             wrt.dir = write_dir,
#             node = "Operations",
#             parm = list(67, 'Line'),
#             value = new.action,
#             overwrite = TRUE)
# 
# inspect_apsimx("consisten_rate_trials-edited.apsimx",
#                node = "Operations",
#                parm = list(67, 'Date'))
# 
# inspect_apsimx("consisten_rate_trials-edited.apsimx",
#                node = "Operations",
#                parm = list(67, 'Action'))
# 
# inspect_apsimx("consisten_rate_trials-edited.apsimx",
#                node = "Operations",
#                parm = list(67, 'Line'))


#######============== Run simulations =============#######
simulation <- apsimx("consisten_rate_trials-edited.apsimx", src.dir = write_dir, value = "Report") %>% 
  data.table()
  
#######============== Clean simulation results =============#######
  simulation_data <- simulation %>%  
  .[, applied_N := as.character(N)] %>% 
  setnames(c("Soil.Nutrient.MineralN(1)", "Soil.Nutrient.MineralN(2)", "Soil.Nutrient.MineralN(3)", "Soil.Nutrient.MineralN(4)", "Soil.Nutrient.MineralN(5)", "Soil.Nutrient.MineralN(6)", "Soil.Nutrient.MineralN(7)", "Soil.Nutrient.MineralN(8)"), c("Mineral_1", "Mineral_2", "Mineral_3","Mineral_4","Mineral_5","Mineral_6","Mineral_7", "Mineral_8")) %>% 
  .[, .(Date, applied_N, NO3, NO3Leaching, AnnualNleaching, AnnualSoilMineral, AnnualNfertilizer, TotalMineralization, SoilMineralization, CropYield, Mineral_1, Mineral_2, Mineral_3, Mineral_4, Mineral_5, Mineral_6, Mineral_7, Mineral_8)] %>% 
  .[, year := as.numeric(format(simulation$Date, "%Y"))] %>%
  .[, month := as.numeric(format(simulation$Date, "%m"))] %>% 
  .[, day := as.numeric(format(simulation$Date, "%d"))]


###--- store the data ---###

simulation_wet_weather <- rbind(simulation_wet_weather, simulation_data)

}

# saveRDS(simulation_result, "/Users/qianqiandu/Desktop/N_Residual_APSIM/Git_Controled/consistent_rate_trials/same_weather_cons_N.rds")




```


```{r}

##-------------------- plot yield ~ applied_N --------------------##

###------ Fake weather yield ------###
yield <- simulation_wet_weather %>% 
  data.table() %>% 
  .[, .(yield = max(CropYield)), by = c("year", "applied_N")]

yield$applied_N <- as.numeric(yield$applied_N)


fig_fake_weather_wet <- ggplot(data = yield)+
  geom_point(aes(x = applied_N, y = yield, color = as.factor(applied_N)))+
  geom_line(aes(x = applied_N, y = yield, color = as.factor(applied_N)))+
  ggtitle("this used fake weather for consistent trials")


###------ real weather yield ------###
yield_real_weather <- simulation_real_weather %>% 
  data.table() %>% 
  .[, .(yield = max(CropYield)), by = c("year", "applied_N")]

yield_real_weather$applied_N <- as.numeric(yield_real_weather$applied_N)


fig_real_weather <- ggplot(data = yield_real_weather)+
  geom_point(aes(x = applied_N, y = yield, color = as.factor(year)))+
  geom_line(aes(x = applied_N, y = yield, color = as.factor(year)))+
  ggtitle("this used real weather")+
  theme_bw()

ggsave(fig_fake_weather, file = "yield_fake_weather.pdf")

```

What doesn't make sense is that 2015 is completely flat, but I used exactly the same weather (2019), so it is not due to the weather. Maybe because of the initial year, I have some parameters missed for the initial point. 


```{r}

NO3 <- simulation_result %>% 
  data.table() %>% 
  .[, .(yield = max(CropYield), NO3), by = c("year", "applied_N")]

ggplot(data = simulation_result)+
  geom_point(aes(x = Date, y = NO3, color = factor(applied_N)))

```







