---
title: "weather_ofpe"
author: "Qianqian Du"
date: "2023-12-14"
output: html_document
---
# Objective:
+ plot the weather, check the correlation between rain and NO3
+ Run OFPE to see how N application rates changes the N residuals

# uplode the packages
```{r, message=FALSE, warning=FALSE}
library(here)
library(apsimx)
library(tidyverse)
library(ggplot2)
library(data.table)
library(oce)
library(lubridate)
library(readxl)
``` 

```{r}
library(knitr)
opts_chunk$set(error = TRUE)
```

```{r}
#look at it with weather
extd.dir = "/Users/qianqiandu/Library/CloudStorage/Box-Box/N_Residual_APSIM/Git_Controled/weather_ofpe"

weather_IA <- read_apsim_met("Iowa_CC.met", src.dir = extd.dir) %>% 
  data.table() %>% 
  mutate(month = case_when(day <= 31 ~ "Jan",
                           day > 31 & day <= 59 ~ "Feb",
                           day > 59 & day <= 90 ~ "Mar",
                           day > 90 & day <= 120 ~ "Apr",
                           day > 120 & day <= 151 ~ "May",
                           day > 151 & day <= 181 ~ "Jun",
                           day > 181 & day <= 211 ~ "Jul",
                           day > 211 & day <= 240 ~ "Aug",
                           day > 240 & day <= 271 ~ "Sep",
                           day > 271 & day <= 301 ~ "Oct",
                           day > 301 & day <= 331 ~ "Nov",
                           day > 331 & day <= 366 ~ "Dec",
                           ))

weather_IA$year <- as.character(weather_IA$year)
weather_IA$month <- as.character(weather_IA$month)

head(weather_IA)
tail(weather_IA)

  
weather_IA_analysis <- weather_IA %>% 
  .[year >= 1990 & year <= 2020, ] %>%   
  .[, .(sum_year = sum(rain)), by = c("year")]


ggplot(data = weather_IA_analysis)+
  geom_point(aes(x = year, y = sum_year))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 90))

```

---

#OFPE

```{r}

data <- read_excel("/Users/qianqiandu/Desktop/12_Nitrogen_and_Carbon_Cycling/Nitrogen_balance.xlsx") %>% data.table() %>% 
  .[SimulationID == 4 |SimulationID == 5 | SimulationID == 6 |SimulationID == 7 | SimulationID == 8, ]

data$SimulationName %>% unique()

data$Clock.Today <- ymd(data$Clock.Today )

data <- data %>% 
  .[, year := year(Clock.Today)] %>% 
  .[, month := month(Clock.Today)] %>% 
  .[, day := day(Clock.Today)]
  

ggplot(data = data)+
  geom_point(aes(x = Clock.Today, y = NO3, color = factor(year)), size = 0.5)+
  theme_bw()

```
The N residuals are not carried over to the next year, this is a problem if running simulations in this way. 

So I need to figure out a way to run APSIM year by year.

# Try to change the NO3 value in the soil at the initial state for each year

```{r}

data[month == 12 & day ==31, .(year, month, day, SimulationID, NO3)]

data <- read_excel("/Users/qianqiandu/Desktop/try/Nitrogen_balance.xlsx") %>% 
  data.table()

data$SimulationName %>% unique()

data$Clock.Today <- ymd(data$Clock.Today )

data <- data %>% 
  .[, year := year(Clock.Today)] %>% 
  .[, month := month(Clock.Today)] %>% 
  .[, day := day(Clock.Today)]
  

ggplot(data = data[year == 2010, ])+
  geom_point(aes(x = Clock.Today, y = NO3, color = factor(year)), size = 0.5)+
  theme_bw()

```

---

# Run for one plot from 2010 to 2015 

Steps for one iteration:

+ run 2010 simulation. 
  + extract the NO3 on Dec 31st 2010.
  
+ run 2011 simulation
  + change the time from 1/1/2021 to 12/31/2021
  + change the N rate as a random number in c(150, 180, 200, 220, 240)
  + change the initial NO3 on 1/1/2021 as the NO3 level on 12/31/2021.
  
+ 2012 do the same thing and continue. 
  
```{r}

results <- data.table()

inspect_apsimx("Nitrogen_residual.apsimx", src.dir = extd.dir,
               node = "Manager", parm = list("Fertilise on fixed date", NA))

#######============== Run simulations =============#######
sim_2010 <- apsimx("Nitrogen_residual.apsimx", src.dir = extd.dir, value = "Report") %>% 
  data.table()

sim_analysis <- sim_2010 %>% 
  .[, year := year(Date)] %>% 
  .[, month := month(Date)] %>% 
  .[, day := day(Date)]


results <- rbind(results, sim_analysis)

#######============== Run simulations for the second year =============#######

# Edit the time for the second year simulation 
edit_apsimx("Nitrogen_residual.apsimx", src.dir = extd.dir,
              wrt.dir = extd.dir, node = "Clock",
              parm = c("Start", "End"), value = c("01/01/2011", "12/31/2011") , 
              verbose = TRUE)

inspect_apsimx("Nitrogen_residual-edited.apsimx", src.dir = extd.dir, node = "Clock")

##--- Or store it in an object for later editing ---##
# pp <- inspect_apsimx("Nitrogen_residual-edited.apsimx", src.dir = extd.dir,
#               node = "Manager", parm = c("Fertilise on fixed dates", 3),
#               print.path = TRUE)

# edit_apsimx("Nitrogen_residual-edited.apsimx", 
#            src.dir = extd.dir, wrt.dir = extd.dir,
#            node = "Other", parm.path = pp, value = 200, 
#            edit.tag = "-pp"
#            )

#--- Edit the fertilizer amount for the second year simulation ---#

# Create a vector of 5 numbers for N rates
numbers <- c(150, 180, 200, 220, 240)

# Randomly select one N rate
random_number <- sample(numbers, 1)

print(random_number)


edit_apsimx("Nitrogen_residual-edited.apsimx", 
           src.dir = extd.dir, wrt.dir = extd.dir,
           node = "Manager", manager.child = "Fertilise on fixed date", 
           parm = "Amount", value = random_number
           )

inspect_apsimx("Nitrogen_residual-edited-edited.apsimx", src.dir = extd.dir,
               node = "Manager", parm = list("Fertilise on fixed date", NA))


#--- adjust the N residual amount based on last year's NO3 value on Dec 31st as the initial point ---#

inspect_apsimx("Nitrogen_residual-edited-edited.apsimx", 
            src.dir = extd.dir,
            node = "Soil",
            soil.child = "Chemical")
# Where did NO3 go?????

# edit_apsim("Nitrogen_residual-edited-edited.apsimx", 
#             src.dir = extd.dir,
#             wrt.dir = ".",
#             node = "Soil",
#             parm = "NO3",
#             value = 200)


# Try another file, which is offered in the APSIM example
extd.dir3 = "/Users/qianqiandu/Desktop/14_Optimum_N_rate"

inspect_apsimx("optimum_N_rate.apsimx", src.dir = extd.dir3,
            node = "Soil",
            soil.child = "Chemical")
# NO3 is there, but can't use the normal way to change the NO3 values. 

edit_apsimx("optimum_N_rate.apsimx", 
            src.dir = extd.dir3,
            wrt.dir = ".",
            node = "Soil",
            soil.child = "Chemical",
            parm = "NO3",
            value = c(200, 0, 0, 0, 0, 0, 0, 0))


#######============== visualize the results =============#######

# ggplot(data = sim_2010_analysis)+
#   geom_point(aes(x = Date, y = NO3))
#   
```

But, `edit_apsimx_replace_soil_profile` will lead to a different story. 

# What's next:

+ NO3 at the initial point is a parameter, instead of a variable. So changing it is a bit more complex than I expected. Figure out how to change the initial NO3 on 1/1/2021 using the NO3 level on 12/31/2021. (But I am not sure if this is the best way of doing this).

+ Double check the purpose of running this simulation. 
  + Run OFPE to see how N application rates changes the N residuals, and how it affect yield? 
  
+ If I can figure the initial NO3 out and get the result. How should I visualize it? 

---



















