---
title: "ofpe_true_yrp"
author: "Qianqian Du"
date: "2024-04-01"
output: html_document
---

Notes: 

+ From the previous week, we realized that in the figure from the slides, we are assuming all the points having the same initial N residuals. However, from the ofpe_trials.rmd simulations, we saw that the estimated yield responses are looking the same over years, which could because the underling initial N are different. So, I got the true true yield response function last week, from consistent trial designs, results were demonstrated in true_yield_response.rmd. 

+ In today's meeting, we talked about the reason why we are seeing the estimated ofpe yield response was different from other true yield responses, is because, the data we used for estimating yield response are from different data generating processing (they are from different yield response functions). 

Objective:

+ Get the true true yield response functions for each plot, using ofpe for the previous years. 
+ When ploting the figure, one way to do it is by different location; another way to do it is to define the categories of the initial NO3. 
+ Estimate the economic loss from running ofpe without controlling NO3. 


# uplode the packages
```{r, message=FALSE, warning=FALSE}

# update the apsimx package as the latest version
# library(remotes)
# remotes::install_github('femiguez/apsimx')

library(apsimx)
library(tidyverse)
library(ggplot2)
library(data.table)
library(oce)
library(lubridate)

``` 

```{r}

### ============ set path to the file ============ ###
extd.dir = "/Users/qianqiandu/Desktop/N_Residual_APSIM/Git_Controled/ofpe_true_yrp"

write_dir = "/Users/qianqiandu/Desktop/N_Residual_APSIM/Git_Controled/ofpe_true_yrp"


###========== Check parameters of the file =========###
inspect_apsimx("ofpe_true_yrp.apsimx", src.dir = extd.dir, node = "Clock")


edit_apsimx("ofpe_true_yrp.apsimx", src.dir = extd.dir,
              wrt.dir = write_dir, node = "Clock",
              parm = "Start", value = "01/01/2015")

edit_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir,
              wrt.dir = write_dir, node = "Clock",
              parm = "End", value = "12/31/2022", overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir, node = "Clock")


inspect_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir, node = "Weather")

edit_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir,
              wrt.dir = write_dir, node = "Weather",
              value = "fake_weather.met", overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir, node = "Weather")

```


```{r}

# final_result <- data.frame()
# final_result_2 <- data.frame()
# final_result_3 <- data.frame()
# final_result_4 <- data.frame()
# final_result_5 <- data.frame()
# final_result_6 <- data.frame()
# final_result_7 <- data.frame()
# final_result_8 <- data.frame()
# final_result_9 <- data.frame()
final_result_10 <- data.frame()


N_rates <- data.frame(
    var_N = as.character(seq(0, 270, by = 30))
  )

number_plot <- 200

applied_N_by_plots <- data.frame(
  year_1 = sample(N_rates$var_N, number_plot, replace = TRUE),
  year_2 = sample(N_rates$var_N, number_plot, replace = TRUE),
  year_3 = sample(N_rates$var_N, number_plot, replace = TRUE),
  plot_id = seq(1, 200, by = 1)) %>% 
  as.list()



# for (j in 1:length(applied_N_by_plots[[3]])) {
for (j in 181:200) {  
  
new.date <- "2015-05-05"
# new.action <- '[Fertiliser].Apply(Amount: 0, Type: Fertiliser.Types.UAN_N, Depth: 30)'
new.action <- paste0("[Fertiliser].Apply(Amount: ",applied_N[[1]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")

# new.line <- '2015-05-05 [Fertiliser].Apply(Amount: 0, Type: Fertiliser.Types.UAN_N, Depth: 30)'
new_line <- paste0("2015-05-05 [Fertiliser].Apply(Amount: ",applied_N[[1]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(3, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(3, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(3, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(3, 'Date'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(3, 'Action'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(3, 'Line'))


##--------------- change the 2nd year's fertilizer amount -------------##
new.date <- "2016-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",applied_N[[2]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2016-05-05 [Fertiliser].Apply(Amount: ",applied_N[[2]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(11, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(11, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(11, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(11, 'Date'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(11, 'Action'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(11, 'Line'))


##--------------- change the 3rd year's fertilizer amount -------------##
new.date <- "2017-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",applied_N[[3]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2017-05-05 [Fertiliser].Apply(Amount: ",applied_N[[3]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(19, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(19, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(19, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(19, 'Date'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(19, 'Action'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(19, 'Line'))


  for (i in 1:nrow(N_rates)) {
    
    ##--------------- change the 4th year's fertilizer amount -------------##
new.date <- "2018-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",N_rates[[1]][i],", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2018-05-05 [Fertiliser].Apply(Amount: ",N_rates[[1]][i],", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(27, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(27, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(27, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(27, 'Date'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(27, 'Action'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(27, 'Line'))


#######============== Run simulations =============#######
simulation <- apsimx("ofpe_true_yrp-edited.apsimx", src.dir = write_dir, value = "Report") %>% 
  data.table()
  

#######============== Clean simulation results =============#######
  simulation_data <- simulation %>%  
  .[, applied_N_i := N_rates[[1]][i]] %>%
  setnames(c("Soil.Nutrient.MineralN(1)", "Soil.Nutrient.MineralN(2)", "Soil.Nutrient.MineralN(3)", "Soil.Nutrient.MineralN(4)", "Soil.Nutrient.MineralN(5)", "Soil.Nutrient.MineralN(6)", "Soil.Nutrient.MineralN(7)", "Soil.Nutrient.MineralN(8)"), c("Mineral_1", "Mineral_2", "Mineral_3","Mineral_4","Mineral_5","Mineral_6","Mineral_7", "Mineral_8"), skip_absent = TRUE) %>% 
  .[, .(Date, NO3, NO3Leaching, AnnualNleaching, AnnualSoilMineral, AnnualNfertilizer, TotalMineralization, SoilMineralization, CropYield, Mineral_1, Mineral_2, Mineral_3, Mineral_4, Mineral_5, Mineral_6, Mineral_7, Mineral_8)] %>% 
  .[, year := as.numeric(format(simulation$Date, "%Y"))] %>%
  .[, month := as.numeric(format(simulation$Date, "%m"))] %>% 
  .[, day := as.numeric(format(simulation$Date, "%d"))] %>% 
  .[, applied_N := sum(unique(AnnualNfertilizer)), by = year]


final_result_10 <- rbind(final_result_10, simulation_data)

  }

}


# saveRDS(final_result_10, "/Users/qianqiandu/Desktop/N_Residual_APSIM/Git_Controled/ofpe_true_yrp/final_result_10.rds")

```


```{r}

opfe_trials <- readRDS("/Users/qianqiandu/Desktop/N_Residual_APSIM/Git_Controled/OFPE_trials/ofpe_fake_weather.rds")

ofpe_2018 <- opfe_trials %>% 
  data.table() %>% 
  .[year == 2018, ] %>% 
  .[, .(yield = max(CropYield)), by = "applied_N"]

### ======= The first 20 locations =====###
simulation_data <- readRDS("/Users/qianqiandu/Desktop/N_Residual_APSIM/Git_Controled/ofpe_true_yrp/final_result_1.rds") %>% 
  data.table() %>% 
  .[year == 2018, ] %>%
  mutate(plot_id = (row_number() - 1) %/% 3650 + 1) %>% 
  .[, .(yield = max(CropYield), NO3, Date), by = c("plot_id", "applied_N")] %>% 
  .[Date == '2018-05-04', ] %>% 
  mutate(NO3_group = case_when(NO3 < 25 ~ "Low",
                               NO3 >= 25 & NO3 < 30 ~ "Med", 
                               NO3 >= 30 ~ "High"))


quantile(simulation_data$NO3)

###------ plot the yield response by each plot------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data)+
  geom_line(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_line(aes(x = applied_N, y = yield), data = ofpe_2018)

###------ plot the yield response by initial NO3 ------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data)+
  geom_line(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data)

ggplot()+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data)+
  theme_bw()


```

```{r}
### ======= The second 20 locations =====###
simulation_data_2 <- readRDS("/Users/qianqiandu/Desktop/N_Residual_APSIM/Git_Controled/ofpe_true_yrp/final_result_2.rds") %>% 
  data.table() %>% 
  .[year == 2018, ] %>%
  mutate(plot_id = (row_number() - 1) %/% 3650 + 1) %>% 
  .[, .(yield = max(CropYield), NO3, Date), by = c("plot_id", "applied_N")] %>% 
  .[Date == '2018-05-04', ] %>% 
  mutate(NO3_group = case_when(NO3 < 25 ~ "Low",
                               NO3 >= 25 & NO3 < 30 ~ "Med", 
                               NO3 >= 30 ~ "High"))


ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_2)+
  geom_line(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_2)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_line(aes(x = applied_N, y = yield), data = ofpe_2018)


###------ plot the yield response by initial NO3 ------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_2)+
  geom_line(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_2)

ggplot()+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_2)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_2)+
  theme_bw()


```


```{r}
### ======= The third 20 locations =====###
simulation_data_3 <- readRDS("/Users/qianqiandu/Desktop/N_Residual_APSIM/Git_Controled/ofpe_true_yrp/final_result_3.rds") %>% 
  data.table() %>% 
  .[year == 2018, ] %>%
  mutate(plot_id = (row_number() - 1) %/% 3650 + 1) %>% 
  .[, .(yield = max(CropYield), NO3, Date), by = c("plot_id", "applied_N")] %>% 
  .[Date == '2018-05-04', ] %>% 
  mutate(NO3_group = case_when(NO3 < 25 ~ "Low",
                               NO3 >= 25 & NO3 < 30 ~ "Med", 
                               NO3 >= 30 ~ "High"))


ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_3)+
  geom_line(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_3)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_line(aes(x = applied_N, y = yield), data = ofpe_2018)


###------ plot the yield response by initial NO3 ------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_3)+
  geom_line(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_3)

ggplot()+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_3)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_3)+
  theme_bw()

```


```{r}

combine_60 <- rbind(simulation_data, simulation_data_2, simulation_data_3)


ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_3)+
  geom_line(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_3)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_line(aes(x = applied_N, y = yield), data = ofpe_2018)


###------ plot the yield response by initial NO3 ------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine_60)+
  geom_line(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine_60)

ggplot()+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine_60)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine_60)+
  theme_bw()


###------ plot the yield response by initial NO3 and the ofpe yield response------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine_60)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine_60)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_line(aes(x = applied_N, y = yield), data = ofpe_2018)


```







