---
title: "ofpe_true_yrp"
author: "Qianqian Du"
date: "2024-04-01"
output: html_document
---

# Notes: 

+ From the previous week, we realized that in the figure from the slides, we are assuming all the points having the same initial N residuals. However, from the ofpe_trials.rmd simulations, we saw that the estimated yield responses are looking the same over years, which could because the underling initial N are different. So, I got the true true yield response function last week, from consistent trial designs, results were demonstrated in true_yield_response.rmd. 

+ In today's meeting, we talked about the reason why we are seeing the estimated ofpe yield response was different from other true yield responses, is because, the data we used for estimating yield response are from different data generating processing (they are from different yield response functions). 

# Objectives:

+ Get the true true yield response functions for each plot, using ofpe for the previous years. 
+ When plotting the figure, one way to do it is by different location; another way to do it is to define the categories of the initial NO3. 
+ Estimate the economic loss from running ofpe without controlling NO3. 


# uplode the packages
```{r, message=FALSE, warning=FALSE}

# update the apsimx package as the latest version
# library(remotes)
# remotes::install_github('femiguez/apsimx')

library(apsimx)
library(tidyverse)
library(ggplot2)
library(data.table)
library(oce)
library(lubridate)

``` 

# Setup APSIM path and template 
```{r}

### ============ set path to the file ============ ###
extd.dir = "/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/ofpe_true_yrp"

write_dir = "/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/ofpe_true_yrp"


###========== Check parameters of the file =========###
inspect_apsimx("ofpe_true_yrp.apsimx", src.dir = extd.dir, node = "Clock")


edit_apsimx("ofpe_true_yrp.apsimx", src.dir = extd.dir,
              wrt.dir = write_dir, node = "Clock",
              parm = "Start", value = "01/01/2015")

edit_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir,
              wrt.dir = write_dir, node = "Clock",
              parm = "End", value = "12/31/2018", overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir, node = "Clock")


inspect_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir, node = "Weather")

edit_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir,
              wrt.dir = write_dir, node = "Weather",
              value = "fake_weather.met", overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx", src.dir = extd.dir, node = "Weather")

```

# Simulations
```{r}
###---- create the dataset to store the result ----###
final_result <- data.frame()

###---- create N seq for trial designs ----###
N_rates <- data.frame(
    var_N = as.character(seq(0, 270, by = 30))
  )

number_plot <- 200

###--- trial design ---###
applied_N <- data.frame(
  year_1 = sample(N_rates$var_N, number_plot, replace = TRUE),
  year_2 = sample(N_rates$var_N, number_plot, replace = TRUE),
  year_3 = sample(N_rates$var_N, number_plot, replace = TRUE),
  year_4 = sample(N_rates$var_N, number_plot, replace = TRUE),
  plot_id = seq(1, 200, by = 1)) %>% 
  as.list()


#################################################
###====== Loop for running simulations =======###
#################################################

# each loop iteration is one plot, 200 plots, so j = 1:200

# for (j in 1:length(applied_N_by_plots[[1]])) {
for (j in 1:200) {  
  
new.date <- "2015-05-05"
# new.action <- '[Fertiliser].Apply(Amount: 0, Type: Fertiliser.Types.UAN_N, Depth: 30)'
new.action <- paste0("[Fertiliser].Apply(Amount: ",applied_N[[1]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")

# new.line <- '2015-05-05 [Fertiliser].Apply(Amount: 0, Type: Fertiliser.Types.UAN_N, Depth: 30)'
new_line <- paste0("2015-05-05 [Fertiliser].Apply(Amount: ",applied_N[[1]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(3, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(3, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(3, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(3, 'Date'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(3, 'Action'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(3, 'Line'))


##--------------- change the 2nd year's fertilizer amount -------------##
new.date <- "2016-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",applied_N[[2]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2016-05-05 [Fertiliser].Apply(Amount: ",applied_N[[2]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(11, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(11, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(11, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(11, 'Date'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(11, 'Action'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(11, 'Line'))


##--------------- change the 3rd year's fertilizer amount -------------##
new.date <- "2017-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",applied_N[[3]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2017-05-05 [Fertiliser].Apply(Amount: ",applied_N[[3]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")

edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(19, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(19, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(19, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(19, 'Date'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(19, 'Action'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(19, 'Line'))


  # inner loop is to get the true yield response function for each plot, i is the N rate applied for that plot.

  # for (i in 1:nrow(N_rates)) {
    
    ##--------------- change the 4th year's fertilizer amount -------------##

new.date <- "2018-05-05"
new.action <- paste0("[Fertiliser].Apply(Amount: ",applied_N[[4]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")
new_line <- paste0("2018-05-05 [Fertiliser].Apply(Amount: ",applied_N[[4]][j],", Type: Fertiliser.Types.UAN_N, Depth: 30)")


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(27, 'Date'),
            value = new.date,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(27, 'Action'),
            value = new.action,
            overwrite = TRUE)


edit_apsimx("ofpe_true_yrp-edited.apsimx",
            src.dir = write_dir,
            wrt.dir = write_dir,
            node = "Operations",
            parm = list(27, 'Line'),
            value = new.action,
            overwrite = TRUE)

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(27, 'Date'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(27, 'Action'))

inspect_apsimx("ofpe_true_yrp-edited.apsimx",
               node = "Operations",
               parm = list(27, 'Line'))


#######============== Run simulations =============#######
simulation <- apsimx("ofpe_true_yrp-edited.apsimx", src.dir = write_dir, value = "Report") %>% 
  data.table()
  

#######============== Clean simulation results =============#######
  simulation_data <- simulation %>%  
  # .[, applied_N_i := N_rates[[1]][i]] %>%
  setnames(c("Soil.Nutrient.MineralN(1)", "Soil.Nutrient.MineralN(2)", "Soil.Nutrient.MineralN(3)", "Soil.Nutrient.MineralN(4)", "Soil.Nutrient.MineralN(5)", "Soil.Nutrient.MineralN(6)", "Soil.Nutrient.MineralN(7)", "Soil.Nutrient.MineralN(8)"), c("Mineral_1", "Mineral_2", "Mineral_3","Mineral_4","Mineral_5","Mineral_6","Mineral_7", "Mineral_8"), skip_absent = TRUE) %>% 
  .[, .(Date, NO3, NO3Leaching, AnnualNleaching, AnnualSoilMineral, AnnualNfertilizer, TotalMineralization, SoilMineralization, CropYield, Mineral_1, Mineral_2, Mineral_3, Mineral_4, Mineral_5, Mineral_6, Mineral_7, Mineral_8)] %>% 
  .[, year := as.numeric(format(simulation$Date, "%Y"))] %>%
  .[, month := as.numeric(format(simulation$Date, "%m"))] %>% 
  .[, day := as.numeric(format(simulation$Date, "%d"))] %>% 
  .[, plot_id := as.character(j)] %>% 
  .[, applied_N := sum(unique(AnnualNfertilizer)), by = year]


final_result <- rbind(final_result, simulation_data)

  # }

}

```


# This is the result of running ofpe in 2018 to get the estimated yield response
```{r}
# saveRDS(final_result, "/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/ofpe_true_yrp/ofpe_2015_2018.rds")

ofpe_2018 <- readRDS("/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/ofpe_true_yrp/ofpe_2015_2018.rds") %>% 
  data.table() %>% 
  .[year == 2018, ] %>%
  .[, c("applied_N", "plot_id")]

combine$plot_id <- as.character(combine$plot_id)

ofpe_2018 <- left_join(ofpe, combine, by = c("applied_N", "plot_id"))


ggplot()+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_smooth(aes(x = applied_N, y = yield), data = ofpe_2018)

```


# Look at the result of each plot's true yield response

```{r}

# combine <- rbind(simulation_data, simulation_data_2, simulation_data_3, simulation_data_4, simulation_data_5,
#                     simulation_data_6, simulation_data_7,simulation_data_8, simulation_data_9, simulation_data_10)

# saveRDS(combine, "/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/ofpe_true_yrp/simulations_200.rds")

combine <- readRDS("/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/ofpe_true_yrp/simulations_200.rds")


combine$applied_N <- as.numeric(combine$applied_N)
ofpe_2018$applied_N <- as.numeric(ofpe_2018$applied_N)


ggplot()+
  geom_point(aes(x = applied_N, y = yield), data = combine, size = 0.3)+
  geom_line(aes(x = applied_N, y = yield, group = plot_id), data = combine, size = 0.3)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018, color = "blue")+
  geom_smooth(aes(x = applied_N, y = yield), data = ofpe_2018, size = 1.2)+
  theme_bw()


###------ plot the yield response by initial NO3 ------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine)+
  geom_line(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine)

ggplot()+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine)+
  theme_bw()


###------ plot the yield response by initial NO3 and the ofpe yield response------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = combine)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_line(aes(x = applied_N, y = yield), data = ofpe_2018)


```

```{r}

### ======= The first 20 locations =====###
simulation_data <- readRDS("/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/ofpe_true_yrp/final_result_1.rds") %>% 
  data.table() %>% 
  .[year == 2018, ] %>%
  mutate(plot_id = (row_number() - 1) %/% 3650 + 1) %>% 
  .[, .(yield = max(CropYield), NO3, Date), by = c("plot_id", "applied_N")] %>% 
  .[Date == '2018-05-04', ] %>% 
  mutate(NO3_group = case_when(NO3 < 25 ~ "Low",
                               NO3 >= 25 & NO3 < 30 ~ "Med", 
                               NO3 >= 30 ~ "High"))


quantile(simulation_data$NO3)

###------ plot the yield response by each plot------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data)+
  geom_line(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_line(aes(x = applied_N, y = yield), data = ofpe_2018)

###------ plot the yield response by initial NO3 ------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data)+
  geom_line(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data)

ggplot()+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data)+
  theme_bw()


```

```{r}
### ======= The second 20 locations =====###
simulation_data_7 <- readRDS("/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/ofpe_true_yrp/final_result_7.rds") %>% 
  data.table() %>% 
  .[year == 2018, ] %>%
  mutate(plot_id = (row_number() - 1) %/% 3650 + 121) %>% 
  .[, .(yield = max(CropYield), NO3, Date), by = c("plot_id", "applied_N")] %>% 
  .[Date == '2018-05-04', ] %>% 
  mutate(NO3_group = case_when(NO3 < 25 ~ "Low",
                               NO3 >= 25 & NO3 < 30 ~ "Med", 
                               NO3 >= 30 ~ "High"))



ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_2)+
  geom_line(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_2)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_line(aes(x = applied_N, y = yield), data = ofpe_2018)


###------ plot the yield response by initial NO3 ------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_2)+
  geom_line(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_2)

ggplot()+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_2)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_2)+
  theme_bw()


```


```{r}
### ======= The third 20 locations =====###
simulation_data_10 <- readRDS("/Users/qianqiandu/Desktop/N_Residual_APSIM/Codes/ofpe_true_yrp/final_result_10.rds") %>%
  data.table() %>% 
  .[year == 2018, ] %>%
  mutate(plot_id = (row_number() - 1) %/% 3650 + 181) %>% 
  .[, .(yield = max(CropYield), NO3, Date), by = c("plot_id", "applied_N")] %>% 
  .[Date == '2018-05-04', ] %>% 
  mutate(NO3_group = case_when(NO3 < 25 ~ "Low",
                               NO3 >= 25 & NO3 < 30 ~ "Med", 
                               NO3 >= 30 ~ "High"))


ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_3)+
  geom_line(aes(x = applied_N, y = yield, color = factor(plot_id)), data = simulation_data_3)+
  geom_point(aes(x = applied_N, y = yield), data = ofpe_2018)+
  geom_line(aes(x = applied_N, y = yield), data = ofpe_2018)


###------ plot the yield response by initial NO3 ------###

ggplot()+
  geom_point(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_3)+
  geom_line(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_3)

ggplot()+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_3)+
  geom_smooth(aes(x = applied_N, y = yield, color = factor(NO3_group)), data = simulation_data_3)+
  theme_bw()

```









