---
title: "diff_N_amount"
author: "Qianqian Du"
date: "2023-11-25"
output: html_document
---
# Objective:
+ Figure out how to extract $N_s$ and $N_p$ from APSIM using R, then see how vary they are.

+ run by different N amount, for different years.
  + seq(100, 200, by = 20), isolate the impact of weather. 

# uplode the packages
```{r, message=FALSE, warning=FALSE}
library(apsimx)
library(tidyverse)
library(ggplot2)
library(data.table)
library(oce)
library(lubridate)
``` 

```{r}
# path to the file
extd.dir = "/Users/qianqiandu/Library/CloudStorage/Box-Box/N_Residual_APSIM/Git_Controled/Learning_APSIM_codes/12_Nitrogen_and_Carbon_Cycling"

###===== edit parameters ===###
# edit_apsimx("Nitrogen_balance.apsimx", src.dir = extd.dir,
#                  wrt.dir = extd.dir, node = "Manager",
#                  manager.child = "Fertilise on fixed dates",
#                  parm = "Amount", value = 200, verbose = TRUE)

inspect_apsimx("Nitrogen_balance.apsimx", src.dir = extd.dir, node = "Clock")

edit_apsimx("Nitrogen_balance.apsimx", src.dir = extd.dir,
              wrt.dir = extd.dir, node = "Clock",
              parm = "Start", value = "01/01/1990", verbose = TRUE)

edit_apsimx("Nitrogen_balance-edited.apsimx", src.dir = extd.dir,
              wrt.dir = extd.dir, node = "Clock",
              parm = "End", value = "12/31/2020", verbose = TRUE)


## Make sure it worked
inspect_apsimx("Nitrogen_balance-edited-edited.apsimx", src.dir = extd.dir, node = "Clock")

inspect_apsimx("Nitrogen_balance-edited.apsimx", src.dir = extd.dir, 
               node = "Manager", parm = list("Fertilise on fixed date", NA))


#######============== Run simulations =============#######
sim_150 <- apsimx("Nitrogen_balance-edited-edited.apsimx", src.dir = extd.dir, value = "Report") %>% 
  data.table()

#######============== process the results to analysis =============#######

sim_150_data <- sim_150 %>% 
  .[, .(Date, SimulationID, NO3, AnnualResidueMineral, AnnualSoilMineral, AnnualN2O, AnnualNfertilizer, SurfaceN, TotalMineralization, CropYield)] %>% 
  .[, year := as.numeric(format(sim_150$Date, "%Y"))] %>%
  .[, month := as.numeric(format(sim_150$Date, "%m"))] %>% 
  .[, day := as.numeric(format(sim_150$Date, "%d"))]


sim_150_data$year %>% unique()
```

# visualize the results

```{r}
figure_2010 <- ggplot()+
  geom_point(aes(x = Date, y = CropYield, color = "CropYield"), 
             data = sim_150_data[year == "2010" & SimulationID == 1, ]
              )+
  geom_point(aes(x = Date, y = SurfaceN*40, color = "SurfaceN"), 
             data = sim_150_data[year == "2010" & SimulationID == 1, ]
             )+
  geom_point(aes(x = Date, y = NO3*4, color = "NO3"), 
             data = sim_150_data[year == "2010" & SimulationID == 1, ]
             )+
    geom_point(aes(x = Date, y = AnnualResidueMineral*10, color = "AnnualResidueMineral"), 
             data = sim_150_data[year == "2010" & SimulationID == 1, ]
             )+
  geom_vline(xintercept = as.numeric(as.Date("2010-05-05")))+
  geom_vline(xintercept = as.numeric(as.Date("2011-05-05")))+
  geom_vline(xintercept = as.numeric(as.Date("2012-05-05")))+
  scale_y_continuous(
    name = "yield",
    sec.axis = sec_axis(~.*1/40, name="Surface N")
  )+
  theme_bw()

figure_2010

```

```{r, cache=TRUE}

fig_NO3_2010_13 <- ggplot()+
  geom_point(aes(x = Date, y = NO3, color = factor(year)), 
             data = sim_150_data[year %in% c(2010, 2011, 2012, 2013) & SimulationID == 1, ]
             )+
  geom_vline(xintercept = as.numeric(as.Date("2010-05-05")))+
  geom_vline(xintercept = as.numeric(as.Date("2011-05-05")))+
  geom_vline(xintercept = as.numeric(as.Date("2012-05-05")))+
  geom_vline(xintercept = as.numeric(as.Date("2013-05-05")))+
  theme_bw()

fig_NO3_2010_13

# Then, run simulation from 1980 to 2020
fig_NO3 <- ggplot()+
  geom_point(aes(x = Date, y = NO3, color = factor(year)), 
             data = sim_150_data[SimulationID == 1, ]
             )+

  theme_bw()

fig_NO3



#--- download weather data in the field location from daymat ---#
daymet_data <- download_daymet("Ames",
                               lat = 42.0308, 
                               lon = -93.6319, 
                               start = 1990, 
                               end = 2023,
                               path = "/Users/qianqiandu/Desktop",
                               internal = FALSE)

read_csv <- read_csv(("/Users/qianqiandu/Desktop/Ames_1990_2023.csv"), skip = 7) %>% 
  data.table() %>% 
  setnames("prcp (mm/day)", "prcp") %>% 
  setnames('tmax (deg c)', "tmax") %>% 
  setnames('tmin (deg c)', "tmim") %>% 
  .[, week := paste0("week_", sprintf("%02d", ceiling(yday / 7)))] %>% 
  .[, .(year, yday, week, prcp, tmax, tmim)] %>% 
  .[, mean_temp := (tmax + tmim)/2] %>% 
  mutate(month = case_when(yday <= 31 ~ "Jan",
                           yday > 31 & yday <= 59 ~ "Feb",
                           yday > 59 & yday <= 90 ~ "Mar",
                           yday > 90 & yday <= 120 ~ "Apr",
                           yday > 120 & yday <= 151 ~ "May",
                           yday > 151 & yday <= 181 ~ "Jun",
                           yday > 181 & yday <= 211 ~ "Jul",
                           yday > 211 & yday <= 242 ~ "Aug",
                           yday > 242 & yday <= 273 ~ "Sep",
                           yday > 273 & yday <= 303 ~ "Oct",
                           yday > 303 & yday <= 334 ~ "Nov",
                           yday > 334 & yday <= 366 ~ "Dec"))


#------- get weekly average precipitation and temperature data -------#
weather_weekly <- read_csv %>% 
  group_by(year, week) %>% 
  summarise(week_prcp = mean(prcp), week_temp = mean(mean_temp)) %>% 
  data.table()


#------- get yearly average precipitation -------#
weather_result_year <- read_csv %>% 
  data.table() %>% 
  .[, .(month_precip = sum(prcp)), by = c("year")] %>% 
  # .[, .(mean_precip = mean(month_precip)), by = "month"] %>% 
  ##---- convert precip from mm to inch ----##
  .[, year_precip_in := month_precip* 0.0393701]

ggplot(data = weather_result_year)+
  geom_line(aes(x = year, y = year_precip_in))


#--- download weather data in the field location from daymat ---#
daymet_data <- download_daymet("Ames",
                               lat = 42.0308, 
                               lon = -93.6319, 
                               start = 1990, 
                               end = 2023,
                               path = "/Users/qianqiandu/Desktop",
                               internal = FALSE)

read_csv <- read_csv(("/Users/qianqiandu/Desktop/Ames_1990_2023.csv"), skip = 7) %>% 
  data.table() %>% 
  setnames("prcp (mm/day)", "prcp") %>% 
  setnames('tmax (deg c)', "tmax") %>% 
  setnames('tmin (deg c)', "tmim") %>% 
  .[, week := paste0("week_", sprintf("%02d", ceiling(yday / 7)))] %>% 
  .[, .(year, yday, week, prcp, tmax, tmim)] %>% 
  .[, mean_temp := (tmax + tmim)/2] %>% 
  mutate(month = case_when(yday <= 31 ~ "Jan",
                           yday > 31 & yday <= 59 ~ "Feb",
                           yday > 59 & yday <= 90 ~ "Mar",
                           yday > 90 & yday <= 120 ~ "Apr",
                           yday > 120 & yday <= 151 ~ "May",
                           yday > 151 & yday <= 181 ~ "Jun",
                           yday > 181 & yday <= 211 ~ "Jul",
                           yday > 211 & yday <= 242 ~ "Aug",
                           yday > 242 & yday <= 273 ~ "Sep",
                           yday > 273 & yday <= 303 ~ "Oct",
                           yday > 303 & yday <= 334 ~ "Nov",
                           yday > 334 & yday <= 366 ~ "Dec"))



#------- get yearly average precipitation -------#
weather_result_year <- read_csv %>% 
  data.table() %>% 
  .[, .(month_precip = sum(prcp)), by = c("year")] %>% 
  # .[, .(mean_precip = mean(month_precip)), by = "month"] %>% 
  ##---- convert precip from mm to inch ----##
  .[, year_precip_in := month_precip* 0.0393701]

ggplot(data = weather_result_year)+
  geom_line(aes(x = year, y = year_precip_in))


ggplot()+
  geom_point(aes(x = year, y = NO3, color = factor(year)),
             data = sim_150_data[SimulationID == 1 & month == 5 & day == 4, ]
             )+
  geom_line(aes(x = year, y = year_precip_in*5), data = weather_result_year)+
  theme_bw()+
  ggtitle("black line is annual precipitation")
  
  







###========= see the summary of N_s, which is the N amount on May 4th, May 5th is the N application day =====###

N_s <- sim_150_data %>% 
  .[month == 5 & day == 4, NO3]

N_s

summary(N_s)

hist(N_s)

```


# Run one year with a squence of N, so we can see the impact of N fertilizer on N residuals, isolating the weather impact.

```{r, cache=TRUE}
###============ set up the simulation date from 2015 to 2019 ========###
edit_apsimx("Nitrogen_balance.apsimx", src.dir = extd.dir,
              wrt.dir = extd.dir, node = "Clock",
              parm = "Start", value = "01/01/2015", verbose = TRUE)

edit_apsimx("Nitrogen_balance-edited.apsimx", src.dir = extd.dir,
              wrt.dir = extd.dir, node = "Clock",
              parm = "End", value = "12/31/2019", verbose = TRUE)



sim_N_data <- data.frame()

for (i in seq(100, 260, by = 20)) {
  
###===== edit parameters ===###
edit_apsimx("Nitrogen_balance-edited-edited.apsimx", src.dir = extd.dir,
                 wrt.dir = extd.dir, node = "Manager",
                 manager.child = "Fertilise on fixed dates",
                 parm = "Amount", value = i, verbose = TRUE)


#######============== Run simulations =============#######
sim_N <- apsimx("Nitrogen_balance-edited-edited-edited.apsimx", src.dir = extd.dir, value = "Report") %>% 
  data.table()

sim_N_result <- sim_N %>% 
  .[, .(Date, SimulationID, NO3, AnnualResidueMineral, AnnualSoilMineral, AnnualN2O, AnnualNfertilizer, SurfaceN, TotalMineralization, CropYield, AnnualNleaching)] %>% 
  .[, year := as.numeric(format(sim_N$Date, "%Y"))] %>%
  .[, month := as.numeric(format(sim_N$Date, "%m"))] %>% 
  .[, day := as.numeric(format(sim_N$Date, "%d"))]


sim_N_data = rbind(sim_N_data, sim_N_result)

}

sim_N_data$AnnualNfertilizer %>% unique()

```

```{r}

sim_N_2015_fig <- ggplot(data = sim_N_data[year == 2015, ])+
  geom_point(aes(x = Date, y = NO3, color = factor(AnnualNfertilizer)), size = 0.8)+
  theme_bw()

sim_N_2015_fig


sim_N_2015_19_fig <- ggplot(data = sim_N_data)+
  geom_point(aes(x = Date, y = NO3, color = factor(AnnualNfertilizer)), size = 0.8)+
  theme_bw()+
  ggtitle("when the N is zero (orange), it should be the N residual from 01-01 to 04-05")

sim_N_2015_19_fig

```
```{r}
#Look at N leaching

sim_N_leaching_fig <- ggplot(data = sim_N_data)+
  geom_point(aes(x = Date, y = AnnualNleaching, color = factor(AnnualNfertilizer)), size = 0.8)+
  theme_bw()

sim_N_leaching_fig
```

# Thoughts: 

+ Figure out how to extract $N_s$ and $N_p$ from APSIM using R, then see how vary they are.

Based on very simple regressions, I think the results are consistent with what we expected- $N_s$ and $N_p$ have pretty large variations. The mean of $N_s$ is ` r N_s %>% mean()`, so estimating $N_s$ should play a significant role in estimating EONR.

```{r}
N_s
```

Look at the weather from these years, I am expecting to see some patterns between precipitation and N_s. 

However, the simulations I ran above were pretty simple. I used default setting in APSIM, just changed the date and N amount. Next, need to keep exploring the variables, and think about the set up of the simulations more carefully. 

---

Later steps to do: 

+ Find a point's historical daily weather data, finding a function in R to generate its weather distribution.
+ Run APSIM on the hypothetical weather data for that point.
+ Solve the ex-ante and ex-post for maximization.





